---
title: "GR_longitudinal_SES_explore"
author: "Bernice Cheung"
date: "11/3/2022"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo=FALSE,warning=FALSE,message=FALSE}
library(tidyverse)
library(psych)
library(ggplot2)
library(janitor)
library(kableExtra)
library(here)
library(lme4)
```

load data:  
845 goals from 246 participants; 194 variables
```{r}
complete_df_goal <- read.csv(here("inputs", "longitudinal_merged_allTime_w.csv"))
indivDiff_df_sub <- read.csv(here("inputs", "indivDiffDf.csv"))
demo_df_sub <- read.csv(here("inputs", "demoDf.csv"))
sub_df <- read.csv(here("inputs", "wideDf.csv"))
```

transform demographic variable (factors)
```{r}
demo_df_sub <- demo_df_sub %>%
  mutate(gender = recode(gender, `1` = "Male", `2` = "Female"), 
         race = recode(race, `1` = "American Indian or Alaska Native", `2` = "Asian", `3` = "Black or African American", `5` = "White", `6` = NA_character_, `7` = NA_character_, `1,5` = "Mixed", `2,5` = "Mixed", `3,5` = "Mixed"),
         ethnicity = recode(ethnicity, `1` = "Hispanic or Latino", `2` = "Not Hispanic or Latino", `3` = NA_character_), 
         education = recode(education, `1` = "Primary School", `2` = "High School", `3` = "Associate's Degree", `4` = "Bachelor's Degree",`5` = "Master's Degree",`6` = "MBA/JD",`7` = "Doctoral PhD/MD"), 
         marital_status = recode(marital_status, `1` = "Single, never married", `2` = "Married or domestic partnership", `3` = "Widowed", `4` = "Divorced", `5` = "Separated"), 
         income_f = recode(income, `1` = "Less than $25,000", `2` = "$25,000 to $34,999l", `3` = "$35,000 to $49,999", `4` = "$50,000 to $74,999",`5` = "$75,000 to $99,999",`6` = "$100,000 to $149,999",`7` = "$150,000 or more")
         )
```

merge individual level data
```{r}
indiv_df <- indivDiff_df_sub %>%
  left_join(demo_df_sub, by = "MTurkCode") %>%
  left_join(select(sub_df, MTurkCode, listNum), by = "MTurkCode")
```



# Descriptive of demographic information

age
```{r}
demo_df_sub %>%
  ggplot() + 
  geom_histogram(aes(x = age), fill = 'darkorange', color = 'black') + 
  geom_vline(aes(xintercept=mean(age)),  linetype="dashed")
  
```

gender
```{r}
library(RColorBrewer)
my_palette <- brewer.pal(7, "Spectral")

demo_df_sub %>%
  ggplot() + 
  geom_bar(aes(x = gender, fill = gender)) + 
  scale_fill_manual(values = my_palette)
```

race
```{r}

demo_df_sub %>%
  ggplot() + 
  geom_bar(aes(x = race, fill = race)) + 
  scale_fill_manual(values = my_palette) + 
  theme(axis.text.x = element_text(angle=45))
```

education
```{r}
demo_df_sub %>%
  ggplot() + 
  geom_bar(aes(x = education, fill = education)) + 
  scale_fill_manual(values = my_palette) + 
  theme(axis.text.x = element_text(angle=45))
```

annual household income
```{r}
demo_df_sub %>%
  ggplot() + 
  geom_bar(aes(x = income_f, fill = income_f)) + 
  scale_fill_manual(values = my_palette) + 
  theme(axis.text.x = element_text(angle=45))
  
```

subjective SES
```{r}
demo_df_sub %>%
  ggplot() + 
  geom_histogram(aes(x = subjectiveSES), fill = "darkorange") + 
  scale_x_continuous(breaks=seq(0, 10, by = 1)) + 
  geom_vline(aes(xintercept=mean(subjectiveSES)),  linetype="dashed")
```

# Subjective SES & Goal Characteristics

## Number of goals

```{r}
ggplot(data = indiv_df,
       mapping = aes(x = subjectiveSES, 
                     y = listNum,
                     color = gender)) +
  geom_point(alpha = .3,
             size = 3) +
  geom_smooth(method = "lm", 
              se = FALSE, 
              size = 1.5) +
  ylim(1,5) + 
  scale_x_continuous(breaks = seq(1, 10, 1)) +
  scale_y_continuous(breaks = seq(1, 5, 5)) +
  scale_color_manual(values = c("darkorange", 
                                "cornflowerblue"))
```

```{r}
ml_ses_listNum <- lm(listNum ~ subjectiveSES*gender, data = indiv_df)
summary(ml_ses_listNum)
```

## Goal Content

Aggregate subjective level goal content
```{r}
content_df_sub <- complete_df_goal %>%
  select(MTurkCode, content) %>%
  group_by(MTurkCode) %>%
  summarise(finance_n = sum(content == "finance"), 
            health_n = sum(content == "physical health"))

indiv_df <- indiv_df %>%
  left_join(content_df_sub, by = "MTurkCode") %>%
  mutate(finance_bi = case_when(finance_n > 0 ~ TRUE,
                                TRUE ~ FALSE), 
         health_bi = case_when(health_n > 0 ~ TRUE,
                                TRUE ~ FALSE), 
        finance_perc = finance_n/listNum, 
        health_perc = health_n/listNum)

```

### Financial goals

```{r}
table(indiv_df$finance_bi)
```

```{r}
t.test(subjectiveSES ~ finance_bi, data = indiv_df, var.equal = FALSE)
```

```{r}
ml_ses_finance <- lm(finance_perc ~ subjectiveSES*gender, data = indiv_df)
summary(ml_ses_finance)
```

```{r}

ggplot(data = indiv_df,
       mapping = aes(x = subjectiveSES, 
                     y = finance_perc,
                     color = gender)) +
  geom_point(alpha = .3,
             size = 3) +
  geom_smooth(method = "lm", 
              se = FALSE, 
              size = 1.5) +
  ylim(1,5) + 
  scale_x_continuous(breaks = seq(1, 10, 1)) +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  scale_color_manual(values = c("darkorange", 
                                "cornflowerblue"))
```

### Health goals


```{r}
t.test(subjectiveSES ~ health_bi, data = indiv_df, var.equal = FALSE)
```

```{r}
ml_ses_health <- lm(health_perc ~ subjectiveSES*gender, data = indiv_df)
summary(ml_ses_health)
```

```{r}

ggplot(data = indiv_df,
       mapping = aes(x = subjectiveSES, 
                     y = health_perc,
                     color = gender)) +
  geom_point(alpha = .3,
             size = 3) +
  geom_smooth(method = "lm", 
              se = FALSE, 
              size = 1.5) +
  ylim(1,5) + 
  scale_x_continuous(breaks = seq(1, 10, 1)) +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  scale_color_manual(values = c("darkorange", 
                                "cornflowerblue"))
```

```{r}

```

